From Net-Zero to Net-Negative: Assessing \hydrogen{} and \carbon{} Network Strategies in Europe

### 2023-11-20 report/submissions/journal/paper/paper.tex
further work

\begin{abstract}
    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing.
    A gap exists in understanding how comprehensive carbon management technologies, including carbon capture, transport, use, and storage, can be effectively integrated and interact with new infrastructures such as a hydrogen transportation network.
    To address this, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution.
    Our findings reveal that implementing carbon management technologies can offer systemic flexibility to the European energy system. Specifically, a carbon transport network facilitates point source carbon capture, thereby reducing the necessity for direct air capture installation. The cost reduction of  average annual savings of €xx billion.
    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.

    % Achieving a carbon-neutral European economy requires internationally coordinated efforts across all energy sectors, especially those with hard-to-abate emissions like heavy industry. Against this backdrop, It is critical to assess the need for comprehensive carbon management involving technologies such as carbon capture, transport, use, and storage to develop an effective carbon strategy for climate neutrality.

    % This paper provides the first cost-optimal design for a European energy system that fully incorporates carbon management technologies including a carbon transport network, and considers all emissions-intensive sectors. It optimizes carbon technologies, hydrogen transport and storage, and the design of renewable energy sources.

    % We show that carbon management technologies aiming at net-zero emissions provide systemic flexibility to the European energy system. A carbon network is cost-effective leading to average cost savings of €xx billion per year. The \carbon{} network enables viable point source carbon capture, reducing the need for direct air capture facilities. The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.

\end{abstract}

@@

In net-zero economies, carbon capture is regarded as a crucial technology for offsetting emissions from hard-to-abate sectors and unavoidable emissions from land use and agriculture. Capturing carbon at point sources and from the atmosphere via Direct Air Capture (DAC) presents new opportunities for producing drop-in fuels such as methane, methanol, or carbon-neutral liquid hydrocarbons. This process is known as Carbon Capture and Utilization (CCU). Additionally, storing carbon in geological formations is referred to as Carbon Capture and Storage (CCS). Together with carbon transport, these technologies are grouped under the term carbon management and are part of the so-called New Carbon Economy~\cite{arniehellerNewCarbonEconomy2019}.

The European Union's goal to achieve climate neutrality by 2050 involves the implementation of carbon management strategies as a crucial component. This objective is integrated into a wide range of programs, funding schemes and initiatives~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}.

%~\cite{lockwoodEuropeanStrategyCarbon} - Clean Air Task Force
%~\cite{righettiSitingCarbonDioxide2017} - co2 pipeline are a mature technology, references to US and Canada co2 pipelines used for enhanced oil recovery
%~\cite{friedmannNETZEROGEOSPHERICRETURN} - high learning rate for co2 pipeline costs
%~\cite{NorthernLightsWhat} - Northern Lights project in Norway, 1.5 Mt/a co2 transport capacity, 1.5 Mt/a co2 storage capacity, 2024 start of operation, targeted scale of 5 Mtpa CO2 Seq. by 2030.

For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year. In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization. Recently, policymakers and industry in Europe have been committing to carbon management projects, planning the first infrastructure components, and developing business models for emerging sectors of the economy~\cite{adomaitisEquinorRWEBuild2023,apnewswireGermanyDrawLegislation2023,KohlenstoffKannKlimaschutz2023,OGETESJoin2022,TESHydrogenLife2023}. Business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand. In this context, the need for an international strategy, analogous to the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022}, becomes apparent to coordinate and support deep decarbonization efforts. Important insights to formulate such a strategy can be gained from energy system models that provide a holistic view of the energy system and its technological interactions. To this end, a number of optimization tools have been developed in recent years to model synergies between various carbon management technologies~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}. In contrast to Integrated Assessment Models, these tools account for the spatial distribution of carbon sources and sinks.
% point out that~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a co2 network, but only considers the power sector without co-optimizing renewable deployment.

% A comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway.
However, the models are often limited with regard to both geographical scope and detail. While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.

In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies. The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors. We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, \hydrogen{}, methane, and \carbon{}. We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks. Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent. We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.

@@

In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions. Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy. Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}. The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites. Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps. We show the sum of all energy demands in \ref{fig:total-demand-bar}. The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions. Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html

\begin{figure}
    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
    \label{fig:total-demand-bar}
\end{figure}

Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants. Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}. We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}. Inter-regional biomass transport is permitted with transport costs considered.
Hydrogen can be gained from electrolysis and steam methane reforming (SMR). Salt caverns for \hydrogen{} storage are derived from~\cite{caglayanTechnicalPotentialSalt2020}.

\dots

% carbon capture technologies
Carbon capture technologies are represented by a set of augmented technologies with integrated carbon capture. These include Allam Cycle, Gas CHP, SMR, DAC, biomass, biogas and capture of process emissions from industry.

Regarding carbon utilization, our model features three drop-in fuel production technologies: methanation, methanolization, and Fischer-Tropsch synthesis. Processed fuels have no usage restrictions. Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs. Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs. Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
For carbon sequestration, we only consider offshore sites as potential sinks. This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas. Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a.

All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.

% \begin{figure}[h]
%     \centering
%     \includegraphics[width=\linewidth]{baseline/figures/90_nodes/sequestration_map.pdf}
%     \caption{Sequestration map}
%     \label{fig:sequestration_map}
% \end{figure}

@@

% hight level trends
- cost reduction with carbon network is significant (2.5\%)

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies. While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both. "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification. The
    }
    \label{fig:cost_area}
\end{figure}

%
\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
    \caption{Share of facilities with integrated carbon capture as a function of emission targets for system with \hydrogen{} and \carbon{} network.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:captureshare_line}
\end{figure}

>>>>

\begin{abstract}
    The transition to a carbon-neutral European economy is a pressing challenge that demands coordinated action across various energy sectors, particularly in emissions-intensive industries like heavy manufacturing. While management of both \hydrogen{} and \carbon{} is considered a critical component of this transition, a gap exists in understanding how new hydrogen infrastructures effectively interact with comprehensive carbon management technologies, including carbon capture, transport, use, and storage. In particular, the role of a carbon transport network as a complementing or substituting infrastructure to hydrogen transport remains unclear. To address this gap, our study employs optimization techniques to design the first cost-optimal European energy system that fully incorporates carbon management technologies, hydrogen transport and storage, and renewable energy sources with a high spatial and temporal resolution considering all energy intensive sectors.
    Our findings reveal that in a either-or scenario, the hydrogen network is superior to the carbon network as it enables the transportation cheap hydrogen inland to serve industrial processes and carbon capture and utilization at point sources. A carbon network facilitates point-source carbon capture and reduces the need for direct air capture, but leaves the scarcity of hydrogen in energy-intensive regions such as industrial clusters.
    However, in a hybrid scenario, the carbon network is a cost-effective complement to the hydrogen network facilitating carbon capture at point sources and transport to sequestration sites.
    We show that theses findings holds true against the backdrop of a net-zero energy system, as well as scenarios where the net-negative emissions are 5\% and 10\% lower than the 1990 baseline levels.
    Overall, our work demonstrates the cost-effectiveness of a multi-grid system that includes both hydrogen and carbon transport networks and power grid expansion to achieve climate neutrality in Europe.
    The paper underpins the need for a complementing hydrogen network and \carbon{} network to achieve climate neutrality.
\end{abstract}

@@

An integrated management of both hydrogen and carbon is considered a crucial component to achieve climate neutrality in energy systems including hard-to-abate sectors. Hydrogen offers an efficient way to transport and store energy from different sources. Carbon, on the other hand, can be effectively captured from industrial processes and the combustion of both fossil and synthetic fuels through Carbon Capture (CC) techniques or harvested directly from the atmosphere using Direct-Air Capture (DAC). Additionally, it can be stored in geological formations, a process known as carbon sequestration or carbon capture and storage (CCS). In combination, both together build the basis for climate-neutral fuels like methane, methanol or Fischer-Tropsch fuels, central to Power-to-X (PtX) and Carbon Capture and Utilization (CCU) strategies.

Recently, policymakers and industry in Europe have been committing to both hydrogen and carbon management projects, planning infrastructure components, and developing business models for emerging sectors of the economy. With the European Union's goal of achieving climate neutrality by 2050, a wide range of programs, funding models and initiatives have been established in this regard~\cite{eu2023netzero,europeangreendeal,europeaninnovationfund}. Initiatives like the European Hydrogen Backbone~\cite{gasforclimateEuropeanHydrogenBackbone2022} or the Hydrogen infrastructure Map~\cite{H2InfrastructureMap} showcase the potential of hydrogen as a fuel and energy carrier. At the same time, business models from companies like Tree Energy Solutions~\cite{TESHydrogenLife2023}, Carbfix~\cite{WeTurnCO2}, and Equinor~\cite{adomaitisEquinorRWEBuild2023} advertise carbon management hubs that provide green hydrogen, methane, or synfuels on the one hand and offer \carbon{} offtake on the other hand. For Europe, the Capture Map~\cite{ToolsGreenTransition} estimates a potential of 1.4~Gt of carbon capture from point sources per year. In combination with large sequestration potentials as stated in~\cite{weiProposedGlobalLayout2021}, this highlights the vast potential for decarbonization. The Northern Lights project in Norway~\cite{NorthernLightsWhat} is planning with a 1.5 Mt/a co2 transport and sequestration capacity to be operating in 2024, expanding to a targeted scale of 5 Mt/a sequestration by 2030.
To this end, the Clean Air Task Force underlines the importance of a carbon transport system in Europe to facilitate the carbon economy~\cite{lockwoodEuropeanStrategyCarbon}.

However, up to this point, it remains unclear how the two management systems of hydrogen and carbon may complement or replace each other, particularly when it comes to their transport systems. A hydrogen network facilitates the transport of energy from renewable sources to regions with high energy demand, also enabling Carbon Capture and Utilization (CCU) from specific point sources. On the other hand, the carbon transport system allows for capturing and transporting carbon from areas with high emissions to regions with significant sequestration potential and abundant renewable sources, thereby enabling cost-effective CCU.
In the literature, the two network approaches and underlying technologies have been discussed in a number of publications, all of which, however, dealt with the isolated effects~\cite{bakkenLinearModelsOptimization2008,morbeeOptimisedDeploymentEuropean2012,stewartFeasibilityEuropeanwideIntegrated2014,oeiModelingCarbonCapture2014,elahiMultiperiodLeastCost2014,burandtDecarbonizingChinaEnergy2019,middletonSimCCSOpensourceTool2020,bjerketvedtOptimalDesignCost2020,weiProposedGlobalLayout2021,damoreOptimalDesignEuropean2021,becattiniCarbonDioxideCapture2022}. In contrast to Integrated Assessment Models, the underlying optimization tools account for the spatial distribution of carbon sources and sinks which are crucial provide a holistic view of the energy system and its technological interactions.. The work in~\cite{neumannBenefitsHydrogenNetwork2022} examines the effect of a hydrogen network in Europe... (extend on hydrogen literature)
The publication by Morbee et al.~\cite{morbeeOptimisedDeploymentEuropean2012} optimizes the topology and capacity of a \carbon{} network in Europe, but only considers the power sector without co-optimizing renewable deployment.
Another comprehensive example is found in~\cite{becattiniCarbonDioxideCapture2022}, which presents a mixed-integer model to optimize the time-evolution of a \carbon{} transport system in Switzerland, connecting to a remote sequestration site in Norway. To this end, it is often argued that \carbon{} pipelines are a mature technology with an expected high learning rate, given the wide-spread installations in US and Canada for enhanced oil recovery~\cite{righettiSitingCarbonDioxide2017,friedmannNETZEROGEOSPHERICRETURN}.
% However, the models are often limited with regard to both geographical scope and detail. While representing a single country with spatial resolution may neglect synergies of international cooperation, a coarse grained representation of multiple countries may neglect important geographical properties.

To our knowledge, no study has yet considered the co-optimization and comprehensive assessment of both \hydrogen{} and \carbon{} networks in a fully sector-coupled energy system. However, we argue that such an assessment is strongly needed to avoid suboptimal investments and to identify synergies between hydrogen and carbon management technologies. In this paper, we present a detailed study of the European energy system for 2050, which includes high geographical resolution and a comprehensive representation of carbon management technologies. The study is conducted using the PyPSA-Eur energy system model and encompasses all relevant energy sectors. We enable the system to optimize the design of renewable energy sources and storage technologies, as well as the transmission of electricity, \hydrogen{}, methane, and \carbon{}. We show that the dynamics of carbon capture, transport, use, and storage in different regions are highly dependent on the availability of renewable energy, \carbon{} pipelines and sequestration sinks. Our evaluation focuses on the transport dynamics of \carbon{} and \hydrogen{} through their respective networks on the European continent. We also analyze how an energy system with limited annual sequestration potential prioritizes decarbonization and fuel switching in various sectors, and how the construction of carbon networks varies based on different levels of available sequestration potential.

@@

\begin{figure}
    \includegraphics[width=\linewidth]{baseline/figures/90_nodes/total-demand-bar.png}
    \caption{Exogenous demand assumptions. The figure shows the total amount of yearly energy demands for each energy carrier and type of end-consumer considered in the model.}
    \label{fig:total-demand-bar}
\end{figure}
%
In our configuration, the model's time horizon spans one year with a temporal resolution of 3 hours and a spatial resolution of 90 regions. Each of the regions consists of a complex subsystem with technologies for supplying, converting, storing and transporting energy. Exogenous assumptions on energy demand and non-abatable emissions are taken from various sources~\cite{piamanzGeoreferencedIndustrialSites2018,muehlenpfordtTimeSeries2019,mantzosJRCIDEES20152018,NationalEmissionsReported2023,EurostatCompleteEnergyBalance,uwekrienDemandlib2023}. The energy demand for electricity, transport, biomass, heat and gas are defined per region and time-step with heat demand being regionally subdivided into shares of urban, rural and industrial sites. Demands for aviation kerosene, methanol for shipping and naphtha for industry are aggregated in the system scope and kept constant throughout all time-steps. We show the sum of all energy demands in \ref{fig:total-demand-bar}. The system emits of 633~Mt \carbon{} per year through non-abatable emissions from industry, aviation, shipping and agriculture into account, 153~Mt of which are fossil-based industrial emissions. Industrial energy demand and excess heat potentials are calculated per node on the basis of~\cite{hotmaps_industrial_db}.
% use https://pypsa-eur.readthedocs.io/en/latest/licenses.html
%
Endogenous assumptions include the expansion of renewable energy sources, storage technologies, and transmission capacities.
The model considers the various energy carriers like electricity, hydrogen, methan, methanol, liquid hydrocarbons and biomass, together with conversions technologies to convert these into each other.
Carbon-neutral electricity is provided by wind, solar, biomass, hydro and nuclear power plants. Weather-dependent power potentials for solar, wind and hydro are calculated from reanalysis and satellite data sets~\cite{hersbachERA5GlobalReanalysis2020,pfeifrothSurfaceRadiationData2017}  per region and time-stamp, using the open-source tool Atlite~\cite{hofmannAtliteLightweightPython2021}.
Solar and wind power can be expanded in alignment with eligible land-use restrictions calculated on the basis of~\cite{eeaCorineLandCover2012,eeaNatura2000Data2016}. We restrict the electric transmission system's expansion to 20\% of its current capacity, acknowledging the challenges in inaugurating new transmission projects.
The regional use of biomass is restricted by its feasible potential derived from~\cite{enspreso_database,instituteforenergyandtransportjointresearchcentreJRCEUTIMESModelBioenergy2015}. Inter-regional biomass transport is permitted with transport costs considered.

Hydrogen can be gained from electrolysis and steam methane reforming (SMR). Geological distribution and potentials for \hydrogen storage in salt caverns are derived from~\cite{caglayanTechnicalPotentialSalt2020}. Re-electrification of hydrogen is possible via fuel cells. If enabled, hydrogen can be transported via pipelines between regions which can be expanded without limit. The topology of the \hydrogen network is mirroring the topology of the electricity network.

Carbon capture (CC) technologies consist of a range of augmented technologies with integrated carbon capture. These include Allam Cycle Power Plants, Gas Combined Heat and Power Plant (CHP), Steam Methan Reforming (SMR), Direc-Air Capture (DAC), biomass combustion, biogas combustion and capture of process emissions from industry.

Our model features three drop-in fuel production technologies for carbon utilization: methanation, methanolization, and Fischer-Tropsch synthesis. The processed fuels are not limited in their total quantity of use or production. Methane is transported through the gas network, while methanol and FT-synfuel can be transferred inter-regionally without additional costs. Synthetic methane substitutes natural gas or biogas, serving combined heat and power plants, residential heating gas boilers, or industrial gas needs. Synthetic methanol decarbonizes marine industry fuel demands, and FT fuels replace fossil oil for naphtha production, aviation kerosene, or agricultural machinery oil.
We only consider offshore sites as potential sinks for carbon sequestration. This choice favors offshore storage due to its typically larger capacity compared to onshore and concerns over public safety for infrastructure near populated areas. Our estimates for storage potential are conservative, limiting total potential to 25~Mt per site, calculating annual storage availability over 25 years, and capping the total sequestered \carbon{} to 200~Mt/a. All technology cost assumptions for 2040 are sourced from an open-source database~\cite{lisazeyenPyPSATechnologydataTechnology2023}.

@@

The comprehensive analysis of the European energy system scenarios in 2050 highlights significant variations in system costs and technology deployment across the different network strategies, as depicted in Figure \ref{fig:cost_bar}. The system cost decomposition for the Baseline, CO2-Net, H2-Net, and Hybrid scenarios are visualized as stacked bar plots, facilitating an insightful comparison of the economic implications of each scenario.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/cost_bar.png}
    \caption[short]{Total annual system cost for the European energy system for the different scenarios subdivided into groups of technologies. While the `Baseline' scenario has neither a \hydrogen{} or \carbon{} network, `Hybrid' is allowed to expand both. "Gas Infrastructure" combines gas facilities for power and heat production, "H$_2$ Infrastructure" combines H$_2$ production, transport and re-electrification. "Carbon Capture at Point Sources" combines all technologies with integrated carbon capture.}
    \label{fig:cost_bar}
\end{figure}

In the Baseline scenario, the absence of dedicated hydrogen and carbon networks necessitates a heavy reliance on local renewable generation and storage solutions, reflected in the substantial investments in solar and wind capacities.
The cost of carbon removal technologies, CC and DAC, also forms a significant portion of the overall expenses. However, this scenario's limitations are apparent, as the lack of transport networks for hydrogen and carbon leads to higher system costs due to inefficiencies and lost opportunities for optimization.

Transitioning to the CO2-Net scenario, we observe a significant cost reduction and a shift in cost allocation.
The implementation of a \carbon{} transport network (\textit{\carbon{}-Net}) facilitates efficient point-source carbon capture, substantially reducing the dependency on direct air capture (DAC) technologies. The improved carbon management infrastructure also allows for a more strategic deployment of renewables, particularly in areas with high sequestration potential, leading to cost savings in renewable generation and storage.

The H2-Net scenario underscores the pivotal role of hydrogen infrastructure. The creation of a hydrogen transport network enables the cost-effective delivery of hydrogen from regions with a surplus of cheap renewable energy to high-demand industrial centers. This scenario shows a pronounced decrease in investments in solar and gas  infrastructure.

The Hybrid scenario represents the most integrated network strategy, combining both CO2 and H2 networks. While the cost-reduction compared to the \hydrogen-Net scenario is marginal, the allocation of costs reveals a propagation of cost savings in DAC facilities.

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/captureshare_line.png}
    \caption{Share of facilities with integrated carbon capture for different system setups for the net-zero emission scenario.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:captureshare_line}
\end{figure}

With the cost reduction across the different system setups, the utilization of carbon capture technologies varies significantly, as depicted in Figure \ref{fig:captureshare_line}. This variation highlights how the implementation of \hydrogen{} and \carbon{} networks unlocks specific carbon capture potentials.

While the share of CC share for the combustion of upgraded biogas (100\%) and Gas in CHP (0\%) remains unchanged across all scenarios, the

\begin{figure}
    \centering
    \includegraphics[width=\linewidth]{comparison/default/figures/90_nodes/energy_balance_bar_carbon.png}
    \caption{Balance of captured carbon across all system setups. Positive values indicate carbon capture, negative values indicate carbon consumption.}
    % A CO2 network unlock BECSS potentials, most biomass emissions are captured at point sources and transported to sequestration sites. cite~\cite{rosaAssessmentCarbonDioxide2021}
    \label{fig:captureshare_line}
\end{figure}

%
